### 2013-10-03 Paper_presentation.tex
Novo rodapé

\documentclass[notes]{beamer}
\usepackage{graphicx}
\usepackage{url}
\usepackage[english, portuguese]{babel}
\usepackage[latin1, utf8]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{fancyhdr}
\mode<presentation>

{
  % A tip: pick a theme you like first, and THEN modify the color theme, and then add math content.
  % Warsaw is the theme selected by default in Beamer's installation sample files.

  %%%%%%%%%%%%%%%%%%%%%%%%%%%% THEME
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
% \usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
 \usetheme{default}
% \usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
% \usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%% COLOR THEME
  %\usecolortheme{albatross}
  %\usecolortheme{beetle}
  %\usecolortheme{crane}
  %\usecolortheme{default}
  %\usecolortheme{dolphin}
  %\usecolortheme{dove}
  %\usecolortheme{fly}
  %\usecolortheme{lily}
  \usecolortheme{orchid}
  %\usecolortheme{rose}
  %\usecolortheme{seagull}
  %\usecolortheme{seahorse}
  %\usecolortheme{sidebartab}
  %\usecolortheme{structure}
  %\usecolortheme{whale}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%% OUTER THEME
  %\useoutertheme{default}
  %\useoutertheme{infolines}
  %\useoutertheme{miniframes}
  %\useoutertheme{shadow}
  %\useoutertheme{sidebar}
  %\useoutertheme{smoothbars}
  %\useoutertheme{smoothtree}
  %\useoutertheme{split}
  %\useoutertheme{tree}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%% INNER THEME
  %\useinnertheme{circles}
  %\useinnertheme{default}
  %\useinnertheme{inmargin}
  %\useinnertheme{rectangles}
  %\useinnertheme{rounded}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \setbeamercovered{transparent} % or whatever (possibly just delete it)
  % To change behavior of \uncover from graying out to totally invisible, can change \setbeamercovered to invisible instead of transparent. apparently there are also 'dynamic' modes that make the amount of graying depend on how long it'll take until the thing is uncovered.

}

% Get rid of nav bar
\beamertemplatenavigationsymbolsempty

% Use short top
\usepackage[headheight=12pt,footheight=12pt]{beamerthemeboxes}
%\addheadboxtemplate{\color{black}}{
%\hskip0.3cm
%\color{white}
\insertshortauthor
%\insertframenumber \ \ \ \ \ \ \ 
%\insertsection \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  \insertsubsection
%\hskip0.3cm}
%\addheadboxtemplate{\color{black}}{
%\color{white}
%\ \ \ \ 
%\insertsection
%}
%\addheadboxtemplate{\color{black}}{
%\color{white}
%\ \ \ \ 
%\insertsubsection
%}
\title{Measurement and Analysis of Child Pornography Trafficking on P2P Networks\nocite{Hurley}}
\subtitle{}

\def\logounicamp{%
\resizebox{!}{7.5ex}{\includegraphics{unicamp.pdf}}
}

\def\logoabaco{%
\resizebox{!}{7.5ex}{\includegraphics{abaco.pdf}}
}

\institute{Instituto de Computação - Unicamp}

\date{\today}

\subject{Talks}

\def\defn#1{{\color{red} #1}}
% Insere o numero da pagina no rodape.
\setbeamertemplate{footline}[text line]{%
  \parbox{\linewidth}{\vspace*{-8pt}\logounicamp\hfill\institute\hfill\inserttitle
  \hfill\logoabaco\hfill\insertpagenumber}}
\setbeamertemplate{navigation symbols}{}

\author{Ryan Hurley, Swagatika Prusty, Hamed Soroush, Robert J. Walls
Jeannie Albrecht, Emmanuel Cecchet, Brian Neil Levine
Marc Liberatore, Brian Lynn, Janis Wolak}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Scheduling}
  \tableofcontents
\end{frame}

\section{Introduction} 
%este é um slide de exemplo
\begin{frame} %começa um novo slide (frame)

\begin{block}{Bloco}

%começa um novo bloco (caixa no slide)

O que eu escrever aqui aparece no bloco.

\end{block} %finaliza o bloco

\begin{block}{Outro bloco}

Posso criar vários blocos em um \textit{frame}...
E fazer com que apareçam um de cada vez...

\end{block} \pause %o comando "\pause" insere uma pausa no slide

\begin{block}{Bloco oculto}

Este bloco só vai aparecer quando eu "clicar" ou pressionar uma tecla.

\end{block}

\end{frame}

\begin{frame} %%[allowframebreaks]
\frametitle{Bibliography}

\bibliographystyle{unsrt}
\bibliography{refs}

\end{frame}

%%%%% Thanks page
\begin{frame}
\frametitle{Thanks}
\vskip20pt

\begin{center}
{\bf \color{alert} Thanks!}
\end{center}

\vskip20pt

\begin{center}

\vskip12pt
\end{center}

\titlepage
\end{frame}

\end{document}