### 2013-10-14 Paper_presentation.tex
More content...

\documentclass[notes]{beamer}
\usepackage{graphicx}
\usepackage{url}
\usepackage[english, portuguese]{babel}
\usepackage[latin1, utf8]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{fancyhdr}
\mode<presentation>

{
  % A tip: pick a theme you like first, and THEN modify the color theme, and then add math content.
  % Warsaw is the theme selected by default in Beamer's installation sample files.

  %%%%%%%%%%%%%%%%%%%%%%%%%%%% THEME
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
% \usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
 \usetheme{default}
% \usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
% \usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%% COLOR THEME
  %\usecolortheme{albatross}
  %\usecolortheme{beetle}
  %\usecolortheme{crane}
  %\usecolortheme{default}
  %\usecolortheme{dolphin}
  %\usecolortheme{dove}
  %\usecolortheme{fly}
  %\usecolortheme{lily}
  \usecolortheme{orchid}
  %\usecolortheme{rose}
  %\usecolortheme{seagull}
  %\usecolortheme{seahorse}
  %\usecolortheme{sidebartab}
  %\usecolortheme{structure}
  %\usecolortheme{whale}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%% OUTER THEME
  %\useoutertheme{default}
  %\useoutertheme{infolines}
  %\useoutertheme{miniframes}
  %\useoutertheme{shadow}
  %\useoutertheme{sidebar}
  %\useoutertheme{smoothbars}
  %\useoutertheme{smoothtree}
  %\useoutertheme{split}
  %\useoutertheme{tree}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%% INNER THEME
  %\useinnertheme{circles}
  %\useinnertheme{default}
  %\useinnertheme{inmargin}
  %\useinnertheme{rectangles}
  %\useinnertheme{rounded}

  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \setbeamercovered{transparent} % or whatever (possibly just delete it)
  % To change behavior of \uncover from graying out to totally invisible, can change \setbeamercovered to invisible instead of transparent. apparently there are also 'dynamic' modes that make the amount of graying depend on how long it'll take until the thing is uncovered.

}

% Get rid of nav bar
\beamertemplatenavigationsymbolsempty

% Use short top
\usepackage[headheight=12pt,footheight=12pt]{beamerthemeboxes}
%\addheadboxtemplate{\color{black}}{
%\hskip0.3cm
%\color{white}
\insertshortauthor
%\insertframenumber \ \ \ \ \ \ \ 
%\insertsection \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  \insertsubsection
%\hskip0.3cm}
%\addheadboxtemplate{\color{black}}{
%\color{white}
%\ \ \ \ 
%\insertsection
%}
%\addheadboxtemplate{\color{black}}{
%\color{white}
%\ \ \ \ 
%\insertsubsection
%}
\title{Measurement and Analysis of Child Pornography Trafficking on P2P Networks\nocite{Hurley}}
\subtitle{}

\def\logounicamp{%
\resizebox{!}{7.5ex}{\includegraphics{unicamp.pdf}}
}

\def\logoabaco{%
\resizebox{!}{7.5ex}{\includegraphics{abaco.pdf}}
}

\institute{Instituto de Computação - Unicamp}

\date{\today}

\subject{Talks}

\def\defn#1{{\color{red} #1}}
% Insere o numero da pagina no rodape.

\setbeamertemplate{headline}[text line]{%
  \parbox{\linewidth}{\vspace*{8pt}\hfill\insertsection}}

\setbeamertemplate{footline}[text line]{%
  \parbox{\linewidth}{\vspace*{-8pt}\logounicamp\hfill\institute\hfill\inserttitle
  \hfill\logoabaco\hfill\insertpagenumber}}
\setbeamertemplate{navigation symbols}{}

\author{Ryan Hurley, Swagatika Prusty, Hamed Soroush, Robert J. Walls
Jeannie Albrecht, Emmanuel Cecchet, Brian Neil Levine
Marc Liberatore, Brian Lynn, Janis Wolak}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Scheduling}
  \tableofcontents
\end{frame}

\section{Introduction} 
%este é um slide de exemplo
\begin{frame} %começa um novo slide (frame)

\end{frame}

\section{Criminal Investigation}
\begin{frame}

\begin{itemize} 
    \item[\checkmark] Works properly and is evaluated under the goal of the investigations.

    \item[\checkmark] We follow that principles, (basic principles) rather than isolated characterization of the users.

    \item[\checkmark] We will review the USA\footnote{Fourth Amendment and related jurisprudence} Law under the constraints of criminal investigations for Children Pornograph.
\end{itemize}

\end{frame}

\begin{frame}
\begin{block}{Works properly and is evaluated under the goal of the investigations}

\begin{itemize}

\item[\checkmark]That  means: The criminal investigation is increasingly advanced and with more development tools for this one.

\item[\checkmark]There are always more groups that works to find and to discover and try to prevent Child Pornography (CP).

\item[\checkmark]But it is very difficult do prevent, because of the large scale growth in the worldwide web. There are over 1,8 milion CP in internet ''found on eMule'' (we estimate much more).

\end{itemize}

\end{block}

\end{frame}

\begin{frame}
\begin{block}{Basic principles rather than isolated characterization of the users}

\begin{itemize}

\item[\checkmark]This means that we will not discuss about a particular user, but the market situation that involves this type of crime.

\item[\checkmark]In criminal investigations of the type we consider search warrants must specify this location, and not a person (not a user).

\item[\checkmark]Actions by the investigators are shortened by law ''Fourth Amendment and Related Jurisprudence'', where this means that the user has a protection on a electronic data.

\end{itemize}

\end{block}

\end{frame}

\begin{frame}
\begin{block}{What is wrong with Fourth Amendment Jurisprudence?}

\textbf{The Third Party Doctrine}

According to the Supreme Court's third party doctrine, personal information, once exposed to any third party, loses all Fourth Amendment protection. Some information exposed to third parties is protected by various statutes, but those can be inconsistent and outdated. The Electronic Communications Privacy Act (ECPA), for example, is notably out of date, leaving privacy protection of technology, as the Ninth Circuit put it, ''a confusing and uncertain area of the law.''. Some privacy interests that are currently unprotected under the Fourth Amendment. Konop ... also receive protection under the First Amendment – but that protection is far from comprehensive... (1967)

\end{block}

\end{frame}

\section{Forensic Measurement}
\begin{frame}

\end{frame}

\section{Availability and Resilience}
\begin{frame}

\end{frame}

\section{FOI Redundancy and Availability}
\begin{frame}

\end{frame}

\section{Comparing Aggressive Peers}
\begin{frame}

\end{frame}

\section{Analysis of User Aliasing}
\begin{frame}

\end{frame}

\section{Measurement Limitations}
\begin{frame}

\end{frame}

\section{Related Work}
\begin{frame}

\end{frame}

\section{Conclusions and Future Work}
\begin{frame}

\end{frame}

\section{References}
\begin{frame} %%[allowframebreaks]

\frametitle{Bibliography}

\bibliographystyle{unsrt}
\bibliography{refs}

\end{frame}

%%%%% Thanks page
\begin{frame}
\frametitle{Thanks}
\vskip20pt

\begin{center}
{\bf \color{alert} Thanks!}
\end{center}

\vskip20pt

\begin{center}

\vskip12pt
\end{center}

\titlepage
\end{frame}

\end{document}